\documentclass[12pt,a4paper]{article}

% ── Encoding & language ───────────────────────────────────────────────
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

% ── Subfiles ──────────────────────────────────────────────────────────
\usepackage{subfiles}

% ── Typography ────────────────────────────────────────────────────────
\usepackage{microtype}
\usepackage[sfdefault]{FiraSans}
\usepackage{setspace}
\doublespacing

% ── Geometry ──────────────────────────────────────────────────────────
\usepackage[top=2.5cm,bottom=2.5cm,left=3cm,right=3cm]{geometry}

% ── Math ──────────────────────────────────────────────────────────────
\usepackage{amsmath}
\usepackage{amssymb}

% ── Tables ────────────────────────────────────────────────────────────
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{array}
\usepackage{rotating}      % sidewaystable
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

% ── Figures ───────────────────────────────────────────────────────────
\usepackage{graphicx}
\graphicspath{{img/}{../img/}}
\usepackage{float}
\usepackage{caption}
\usepackage{subfig}

% ── Hyperlinks & citations ────────────────────────────────────────────
\usepackage[hidelinks,breaklinks]{hyperref}
\usepackage[numbers,sort&compress]{natbib}

% ── Line numbers (review) ─────────────────────────────────────────────
\usepackage{lineno}
\linenumbers

% ── Misc ──────────────────────────────────────────────────────────────
\usepackage{xcolor}
\usepackage{url}

% ═════════════════════════════════════════════════════════════════════
\begin{document}

% ── Title & authors ───────────────────────────────────────────────────
\title{\textbf{Automated Risk of Bias Assessment in Systematic Reviews
Using Large Language Models: A Multi-Model Single-Agent Benchmark
on 227 Randomised Controlled Trials}}

\author{
  Luiz Felipe Guain Teixeira\textsuperscript{1,2,3}\quad
  Cassiano Pereira de Barros\textsuperscript{4}\quad \\
  Renato Vicente\textsuperscript{1,2,3}
  \\[6pt]
  \small\textsuperscript{1}TELUS Digital Research Hub\\
  \small\textsuperscript{2}Center for Artificial Intelligence and Machine Learning, University of São Paulo\\
  \small\textsuperscript{3}Institute of Mathematics, Statistics and Computer Science, University of São Paulo\\[2pt]
  \small\textsuperscript{4}Paulista School of Medicine, Federal University of São Paulo\\[2pt]
  \small \texttt{\{luiz.felipe.teixeira, rvicente\}@usp.br}
}

\date{}
\maketitle
\thispagestyle{empty}

% ── Abstract ──────────────────────────────────────────────────────────
\begin{abstract}
\noindent\textbf{Background:}
Risk of bias (RoB) assessment is a mandatory but labour-intensive step in
systematic review production.
Large language models (LLMs) offer the potential to automate this task in a
zero-shot setting, yet systematic benchmarks spanning multiple frontier models
and inference strategies are lacking.

\noindent\textbf{Methods:}
Eight state-of-the-art LLMs (Gemini~2.5~Pro, Gemini~2.5~Flash, GPT-5.4,
Claude~Haiku~4.5, DeepSeek~v4-Flash, Grok~4.3, Qwen~3.7-Plus, and GLM-5)
were evaluated on zero-shot, single-agent RoB~1.0 assessment of 227~randomised
controlled trials drawn from 127~Cochrane systematic reviews published in~2024.
Models were accessed via the OpenRouter API in June~2026 using a fixed prompt
and provider-default generation parameters (exact OpenRouter identifier strings
in Table~\ref{tab:models}).
Ground-truth annotations were taken from the published Cochrane consensus
judgements across seven methodological domains (1{,}589~cells total).
Performance was assessed using accuracy, Macro-F1, and Gwet's~AC1, with
study-clustered bootstrap 95\% confidence intervals.
Two models (Gemini~2.5~Flash and DeepSeek~v4-Flash) were additionally
evaluated under a single-call inference mode for comparison with the
per-domain strategy.

\noindent\textbf{Results:}
Overall accuracy ranged from 57.6\% to 62.5\% across models, a spread
smaller than the width of individual 95\% confidence intervals;
Macro-F1 confidence intervals overlapped across all eight models.
AC1 provided the clearest separation: Claude~Haiku~4.5
(AC1~$= 0.485$; 95\%~CI $[0.443, 0.526]$) had non-overlapping CIs relative to
Gemini~2.5~Pro ($[0.344, 0.420]$) and GPT-5.4 ($[0.349, 0.426]$),
though this comparison is unadjusted and exploratory.
Unclear Risk was the most consistently under-detected class
(recall~$< 0.20$ in three of seven domains across most models).
Consensus failures (cells where no model was correct) accounted for only
14.2\% of the benchmark, confirming substantial non-redundancy among models;
the remaining 85.8\% of cells had at least one correct model.
The comparison of inference modes yielded small, largely non-significant
differences: ten of twelve $\Delta$ metrics had confidence intervals
spanning zero, and D7~(Other Bias)---the domain hypothesised to benefit
most from joint assessment---showed negligible change under either mode.

\noindent\textbf{Conclusions:}
Frontier LLMs can perform zero-shot RoB~1.0 assessment at accuracy levels
broadly consistent with estimates of human inter-rater variability for this task
\citep{Hartling2013}, though direct comparison on this specific corpus was not
performed; no single model dominates across all metrics, and no consistent
advantage was found for single-call over per-domain inference.
The primary practical distinction between inference modes is computational:
single-call reduces inference calls by 7$\times$ without a reliable accuracy
trade-off.
All analyses are exploratory; results should be interpreted as directional
benchmarks rather than definitive rankings.

\noindent\textbf{Keywords:} risk of bias; systematic review; large language
models; natural language processing; Cochrane; evidence synthesis; automation
\end{abstract}

\newpage

% ── Main sections ─────────────────────────────────────────────────────
\subfile{sections/introduction}

\subfile{sections/methods}

\subfile{sections/results}

\subfile{sections/conclusions}

% ── Acknowledgements ──────────────────────────────────────────────────
\section*{Acknowledgements}
We gratefully acknowledge the financial support of the TELUS Digital Research Hub.

\section*{Data and Code Availability}
The analysis code is publicly available at
\url{https://github.com/ACS-USP/systematic-reviews}.

\section*{Conflicts of Interest}
The authors declare no conflicts of interest.

% ── Bibliography ──────────────────────────────────────────────────────
\bibliographystyle{unsrtnat}
\bibliography{references}

% ── Supporting Information ────────────────────────────────────────────
\newpage
\subfile{sections/suppinfo}

\end{document}
