\documentclass[../main.tex]{subfiles}
\begin{document}

\section{Conclusions}
\label{sec:conclusions}

Eight state-of-the-art LLMs assessed RoB~1.0 for 227 randomised controlled trials
in a zero-shot, single-agent setting, addressing whether the reliability shortfall
documented in prior work is uniform across the RoB~1.0 structure or varies
predictably by domain and intervention type.
Overall accuracy ranged narrowly from 57.6\% to 62.5\%, but Macro-F1 and Gwet's~AC1
revealed meaningful differences in how models handled the three nominal classes.
Models with the highest accuracy achieved it partly by concentrating predictions on
majority-class labels, a strategy that inflates accuracy whilst suppressing detection
of High Risk and Unclear Risk judgements---the very categories most consequential for
evidence appraisal.
Macro-F1 and Gwet's~AC1 are therefore more informative primary metrics than accuracy
for this task.

\emph{Unclear Risk} was the most consistently under-detected class across all models
and domains, with recall below 0.20 in three of seven domains.
This systematic failure reflects the epistemic character of the category---it
requires the model to recognise that key methodological information is absent,
a more complex inference than identifying a positive or negative methodological
signal in the text.

Domain-level and stratification analyses revealed structured heterogeneity in
performance.
Sequence generation~(D1) and allocation concealment~(D2) were reliably assessed
across all models, whereas other sources of bias~(D7) showed the greatest
inter-model accuracy spread and blinding of outcome assessors~(D4) had the
highest consensus failure rate (22.0\%---no model correct in 50 of 227~cells).
Clinical domains with clear, domain-specific methodological criteria (Orthopaedics
\& Trauma, Endocrine \& Metabolic) supported higher Macro-F1 than domains with more
heterogeneous study designs (Infectious Disease, Reproductive \& Sexual Health).

Consensus failures---cells where no model was correct---accounted for 14.2\%
(226 of 1{,}589 cells) and were concentrated in D4 and D5, the two most
ambiguous domains.
These cases identify the boundaries of current frontier model capability under
this evaluation design and may reflect genuine annotation uncertainty as much
as model failure.
The remaining 85.8\% of cells had at least one correct model, yet pairwise
complementarity analysis showed that 7.4\%--17.4\% of cells are correctable
by substituting any one model for another, confirming that the eight models
make substantially non-redundant errors.

The comparison between per-domain and single-call inference modes
(Section~\ref{sec:inference}) produced small differences that do not
collectively support a performance advantage for either strategy.
Ten of twelve $\Delta$ values across both models and six metrics had 95\%~CIs
spanning zero; the two exceptions were class-specific effects of limited
practical magnitude (Unclear Risk F1 for Flash: $-$0.044; Low Risk F1 for
DeepSeek: $+$0.026) that did not propagate to aggregate metrics.
Most notably, D7~(Other Bias)---the domain for which joint contextual
assessment was hypothesised to be most beneficial---showed negligible
differences under both modes ($\pm$1~pp), indicating that its difficulty
is intrinsic to the vagueness of the criterion rather than a consequence
of fragmented inference calls.
Taken together, the preliminary evidence from this two-model comparison does
not identify a meaningful performance rationale for preferring single-call
over per-domain assessment; the primary practical distinction between the
two modes is computational (single-call reduces inference calls by 7$\times$)
rather than one of accuracy.

Reframing High Risk detection as a binary classification task (High Risk
vs.\ not High Risk) highlights the clinical stakes of the leniency bias
documented above.
Under this framing, the best-performing models for High Risk recall detected
fewer than two thirds of High Risk cells (F1-HR: 0.455--0.480 for the
top-ranked models), meaning roughly one in three cells that should trigger
reviewer scrutiny passed undetected.
Because the cost of a false negative---a biased study classified as acceptable
and included unquestioned in a meta-analysis---substantially exceeds the cost of
a false positive, tools intended to assist real-world screening workflows would
need to prioritise sensitivity over overall accuracy, a trade-off not currently
optimised by the zero-shot prompt evaluated here.

The cell-level concordance partition offers a practical framework for directing
qualitative review effort.
Cells in which all models converge on the correct answer require no additional
attention.
Cells in which all models converge on the \emph{same wrong} answer are the
highest-priority targets for prompt refinement or annotation review: unanimity
of error across architecturally diverse models makes model-idiosyncratic
explanation implausible, pointing instead to shared training priors, prompt
ambiguity, or genuine ground-truth noise.
Cells in which a single model is the sole outlier---either the only model
correct or the only model incorrect---index model-specific capacity gaps
independent of task difficulty.
Targeting qualitative analysis at consensus errors before individual outliers
maximises diagnostic yield per unit of human effort.

The domain and stratification analyses converge on a specific, actionable
direction for prompt improvement: a blinding rule sensitive to intervention
type.
Models would benefit from explicit guidance distinguishing trials in which
blinding is physically feasible (pharmacological interventions with matched
placebos) from those in which it is not (surgical procedures, physiotherapy,
medical devices), with the latter receiving a conservative default that does
not presuppose the availability of a blinding mechanism.
This single rule targets D4---the domain with the most systematic and
directionally consistent errors in the benchmark---and has direct implications
for the High Risk recall deficit identified above.

The aggregate performance range observed here is consistent with the recent
preregistered evaluation by \citet{Nyrhi2026}, who applied four contemporary LLMs
to 100~RCTs from Cochrane reviews under RoB~1.0 and reported interobserver
$\kappa$ of 0.27--0.39---concluding, as we do, that current models do not meet the
reliability threshold for autonomous use.
Our study extends that comparison in three directions.
First, evaluating eight models spanning five provider families reveals that aggregate
performance is approximately stable across frontier models, shifting the central
question from \emph{which model} to \emph{which domains and intervention types}
support reliable assistance---a question that aggregate summaries cannot answer.
Second, domain-stratified analysis shows that D1 and D2, where models achieve
accuracy and concordance broadly comparable to a second human reviewer, are
plausible targets for assisted screening today, whereas the systematic leniency
bias in D4 and the idiosyncratic divergence in D7 identify specific failure modes
that require targeted prompt improvements before reliable assistance is achievable.
Third, explicit degenerate baselines and the juxtaposition of Cohen's~$\kappa$
with Gwet's~AC1 expose analytical pitfalls specific to this task: the margin above
trivially achievable strategies is narrowest for the clinically consequential
classes (High Risk and Unclear Risk), and the two agreement coefficients diverge
most sharply for the model with the highest accuracy (Claude~Haiku~4.5:
$\kappa = 0.325$ vs AC1~$= 0.485$), cautioning against treating AC1-based
summaries as a direct proxy for classification performance when class distributions
are skewed.

Several limitations should be noted.
First, the ground-truth annotations derive from published Cochrane review
reports and were not independently re-annotated for this study.
The Cochrane annotations represent the consensus output of a structured
multi-reviewer process (independent assessment by at least two trained
reviewers, with disagreements resolved by discussion or arbitration), which
is the reference standard used operationally by the field and the most
appropriate benchmark for evaluating a tool intended to assist real-world
systematic review workflows.
Quantifying internal reliability within this corpus would require re-annotating
a subsample with independent trained reviewers; however, agreement between
Cochrane-trained reviewers and non-specialist re-annotators would conflate
task difficulty with expertise asymmetry---the same confound that limits the
interpretability of comparisons between Cochrane and external reviewers
\citep{ArmijoOlivo2014}.
The relevant population-level IRR estimate is therefore taken from Hartling
et al.\ \citep{Hartling2013}, who measured agreement between formally trained
reviewer pairs applying the same protocol: $\kappa = 0.79$ for D1 but only
$\kappa = 0.24$--$0.37$ for remaining domains, falling to $\kappa = 0.05$--$0.09$
between independent consensus pairs; independent Cochrane teams show
similarly variable agreement \citep{Konsgen2020}.
These figures establish that the reference standard itself carries substantial
noise, placing a practical ceiling on any automated system's agreement with
published annotations that is independent of model quality.
Converting the domain-level $\kappa$ values for the non-D1 domains
(0.24--0.37) to expected accuracy using the observed pooled class distribution
(LR\,53.5\%, HR\,17.4\%, UR\,29.1\%) yields approximately 58--62\%---a range
that coincides closely with the model accuracy range observed in this study
(57.6\%--62.5\%), suggesting that the evaluated LLMs perform at a level broadly
comparable to a second independent human reviewer across most domains.
Second, each model was queried once per study using provider-default generation
parameters.
This reflects the intended evaluation scope: zero-shot performance as models
are available to practitioners, without artificial parameter tuning.
Manipulating generation temperature would introduce heterogeneous conditions
across model families, and strategies such as majority voting over repeated
draws represent post-hoc optimisation outside the zero-shot benchmark design.
As a consequence, differences between models separated by fewer than five
percentage points should not be interpreted as definitive rank orderings in the
absence of confidence intervals; these are discussed as directional trends.
Third, all models were accessed in June~2026 via the OpenRouter~API using the
identifier strings reported in Table~\ref{tab:models}; providers may silently
update the weights associated with a fixed identifier, so exact reproduction
requires accessing the same model strings within the same inference window.
Results should therefore be interpreted as reflecting the checkpoint state active
in June~2026, not any future revision of the same identifier.
Fourth, the single-agent zero-shot design examines one specific slice of the
capability space; few-shot prompting, retrieval augmentation, fine-tuning, and
multi-agent architectures were not evaluated and may yield substantially
different results.
Fifth, training-data contamination cannot be excluded, but several structural
features of the evaluation design limit its plausibility as a primary driver
of results.
Memorisation-based performance would require a model to (i)~have seen the
particular one of the 127 distinct Cochrane~2024 reviews from which each study
was drawn, (ii)~correctly identify that the input Markdown document corresponds
to a specific entry in that review's RoB table, and
(iii)~reproduce the reviewer's label rather than reason from the text.
The Cochrane reviews sampled here were published in~2024; most included RCTs
appear in paywalled journals, making full-text ingestion at training time less
likely for the majority of papers.
Moreover, the empirical pattern is inconsistent with memorisation at scale:
the best individual model (Claude~Haiku~4.5) achieved a perfect 7/7 score on
only 12 of 227 studies (5.3\%), and even across the full ensemble of eight
models all seven domains were correct in only 40 studies (17.6\%).
Both figures are consistent with the rates expected under independent domain
judgements at the observed per-domain accuracy ($0.60^7 \approx 2.8\%$; inflated
modestly by the low intraclass correlation of $\hat\rho = 0.044$ to roughly
4--5\% per model; see Supporting Information~S7 for derivation), leaving no systematic excess attributable to retrieval.
Targeted contamination probes (e.g.\ membership inference or verbatim
reproduction tests) were not conducted, as the opaque training corpora of the
evaluated models preclude rigorous analysis; this remains a caveat for
interpretation.
Sixth, the Cochrane corpus used here comprises better-reported trials than the
broader trial literature, so these performance estimates may be optimistic
relative to unselected primary research.\end{document}
